---
title: "8-21-25 Notes"
output: html_document
date: "2025-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
data(mtcars)
```
The built-in R dataset, mtcars, has 11 variables:

- mpg (miles per gallon)  
- cyl (number of cylinders)  
- disp (displacement)  
- hp (horsepower)  
- drat (rear axle ratio)  
- wt (weight)  
- qsec (quarter mile time)  
- vs (engine type)  
- am (transmission type)  
- gear (number of forward gears)  
- carb (number of carburetors)  
    
The purpose of this project is to understand what impacts the quater mile time of cars in `mtcars`.

Below is a summary of the initial data set:
```{r,echo=FALSE}
data(mtcars)
summary(mtcars)
```

The first step of this process is to create a scatter plot for all other variables in relation to `qsec`. The following function was created to more efficiently create the graphs.
```{r}
#Plot qsec vs any variable
plot_mtcars <- function(var) {
  x <- mtcars[[var]]
  y <- mtcars$qsec
  
  plot(x, y,
       xlab = var,
       ylab = "Quarter Mile Time (qsec)",
       main = paste("Quarter Mile Time vs", var),)
}
```

Using that function all 10 plots were created. Examining the plots the quater mile time seemed related too:

``` {r, echo=FALSE}
plot_mtcars("cyl")
plot_mtcars("disp")
plot_mtcars("hp")
plot_mtcars("vs")
plot_mtcars("am")
plot_mtcars("wt")
```

From these relationships multicollinearity concerns arise from number of cylinders to horsepower and displacement to horsepower. This relationship is shown by plotting the two together.
```{r, echo=FALSE}
plot(mtcars$cyl, mtcars$hp,
     xlab = "Cylinders",
     ylab = "Horsepower",
     main = "Horsepower vs Cylinders",)
plot(mtcars$disp, mtcars$hp,
     xlab = "Displacement",
     ylab = "Horsepower",
     main = "Displacement vs Cylinders",)
```
